---
title: "Assignment 3"
author: "Michaela Rieser"
date: 'Today''s date here: 10/21/2021'
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Collaborators: **.

This assignment is due on Canvas on Wednesday 10/27/2021 before class, at 10:15 am. Include the name of anyone with whom you collaborated at the top of the assignment.

Submit your responses as either an HTML file or a PDF file on Canvas. Also, please upload it to your website.

Save the file (found on Canvas) crime_simple.txt to the same folder as this file (your Rmd file for Assignment 3).

Load the data.
```{r}
library(readr)
library(knitr)
dat.crime <- read_delim("crime_simple.txt", delim = "\t")
```

This is a dataset from a textbook by Brian S. Everitt about crime in the US in 1960. The data originate from the Uniform Crime Report of the FBI and other government sources. The data for 47 states of the USA are given. 

Here is the codebook:

R: Crime rate: # of offenses reported to police per million population

Age: The number of males of age 14-24 per 1000 population

S: Indicator variable for Southern states (0 = No, 1 = Yes)

Ed: Mean of years of schooling x 10 for persons of age 25 or older

Ex0: 1960 per capita expenditure on police by state and local government

Ex1: 1959 per capita expenditure on police by state and local government

LF: Labor force participation rate per 1000 civilian urban males age 14-24

M: The number of males per 1000 females

N: State population size in hundred thousands

NW: The number of non-whites per 1000 population

U1: Unemployment rate of urban males per 1000 of age 14-24

U2: Unemployment rate of urban males per 1000 of age 35-39

W: Median value of transferable goods and assets or family income in tens of $

X: The number of families per 1000 earning below 1/2 the median income


We are interested in checking whether the reported crime rate (# of offenses reported to police per million population) and the average education (mean number of years of schooling for persons of age 25 or older) are related. 


1. How many observations are there in the dataset? To what does each observation correspond?

```{r}
summary(dat.crime)
```

__There are 47 observations in this dataset.  They correspond to 47 US states.__

2. Draw a scatterplot of the two variables. Calculate the correlation between the two variables. Can you come up with an explanation for this relationship?

```{r, fig.width=6, fig.height=4}
reg.output.nc <- lm(formula = R ~ Ed, data = dat.crime)
plot(dat.crime$R,dat.crime$Ed, main = "Relationship between reported crime rate and average education for 47 states", xlab = "Reported Crime rate (# of offenses reported to police per million population)", ylab = "Average education (mean number of years of schooling for persons of age 25 or older)")
abline(reg.output.nc, col="skyblue")
```
```{r}
cor(dat.crime$R, dat.crime$Ed)
```
__Looking at the scatter plot, it does not appear that there is a linear relationship between the reported crime rate and average education.  In addition, the correlation between the two variables is only 0.3228349.  The blue line on the scatter plot shows that there is a positive correlation between the number of offenses reported to police (per million population) and the mean number of years of schooling for persons of ages 25+.  A correlation of 0.3228 corresponds to a weak, positive linear relationship.  Looking at the scatter plot and the correlation value, I would start to conclude that a linear model might not be the best way to to describe the relationship between these two variables.__


3. Regress reported crime rate (y) on average education (x) and call this linear model `crime.lm` and write the summary of the regression by using this code, which makes it look a little nicer `{r, eval=FALSE} kable(summary(crime.lm)$coef, digits = 2)`. (y~x: y regressed on x)
```{r}
crime.lm <-lm(formula=R~Ed, data = dat.crime)
summary(crime.lm)
``` 

```{r} 
kable(summary(crime.lm)$coef, digits = 2)
```

4. Are the four assumptions of linear regression satisfied? To answer this, draw the relevant plots. (Write a maximum of one sentence per assumption.)

```{r} 
plot(dat.crime$Ed, crime.lm$residuals, ylim = c(-50,50), main = "Residuals vs. x", xlab = "x, average years of education", ylab="Residuals" )
abline(h=0, lty="dashed")
```

__To check for error independence (which technically is not possible), we can look at the plot of residuals against x and check for patterns; there are not any patterns in the residuals vs. x graph, so we can assume that the independence assumption is met (assumption2).__  

```{r}
plot(crime.lm, which = 1)
```

__Looking at the residuals vs. fitted plot, the scatter plot smoother line that shows the average value of the residuals at each value of the fitted value, is relatively flat and close to the dashed line, which means that there is no noticeable non-linear trend to the residuals and they appear to be equally variable across the range of fitted values so we can assume that linearity assumption is met (assumption1).__


```{r}
plot(crime.lm, which=3)
```

__The red line is approximately horizontal and there aren't any significant trends which means our residuals have constant variance, so the equal variance  assumption is met (assumption3).__    
```{r}
plot(crime.lm, which = 5)
```
```{r}
plot(crime.lm, which=2)
```

__The Normal QQ plot shows that the tails do not lie along the line as well as they could, but most of the points fall along the line, so we can assume normality (assumption4).__

5. Is the relationship between reported crime and average education statistically significant? Report the estimated coefficient of the slope, the standard error, and the p-value. What does it mean for the relationship to be statistically significant?
```{r}
summary(crime.lm)
```
__The estimated coefficient for the slope for this relationship is 1.1161. The standard error for slope is 0.4878 and the standard error for the intercept is 51.8104.  The p-value for the slope is 0.0269 and it is much larger for the intercept, 0.5996.  The p-value for the intercept is not significant.__

6. How are reported crime and average education related? In other words, for every unit increase in average education, how does reported crime rate change (per million) per state?
__The slope term in this model is saying that for every 1 year of average education increase, the number of offenses reported to police per million population goes up by by 1.1161.  However, we are not saying that there is a direct cause and effect in which when average education increases, the amount of offenses reported to police(per million population) has to also increase.__

7. Can you conclude that if individuals were to receive more education, then crime will be reported more often? Why or why not?

__No, I cannot conclude that if individuals were to receive more education, then crime will be reported more often.  The p-values are not significant enough to reject the null hypothesis.  The null hypothesis states that the slope is zero, meaning that there is no relationship between average education and the number of offenses reported to the police.__


