---
title: "assignment4"
author: "Michaela Rieser"
date: "11/3/2021"
output: html_document
---
# Assignment 4 

```{r}
ggplot2::mpg
```
__This is the mpg data frame for the data set we are looking at.  It is a collection of variables (columns) and observations (rows).  mpg contains observations collected by the US Environmental Protection Agency on 38 models of car.__
```{r}
?mpg
```
__This code tells us more about the data set.__
```{r}
library(ggplot2)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```
__This plot shows the relationship between engine size (displ) and fuel efficiency (hwy).  From this, we can see that there is a negative association between the two variables.  So this implies that when a car has a bigger engine,it uses more fuel. ggplot(data=mpg) creates an empty graph and you have to add more layers with the other code.__

### Aestetic Mappings 
There is a plot in which there are points that fall outside of the linear trend (in red).  Those cars have higher mileage that expected.  A third variable can be added to a scatter plot mapping it to an aesthetic.  Aesthetics include things like the size, the shape, or the color of your points.  Can use the world level to describe aesthetic properties.  
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```
__This new scatter plot added a third variable, class.  By looking at the color of the dot, we can determine the class of each vehicle. The points that do not follow the linear trend are 2 seater vehicles.  To map an aesthetic to a variable, associate the name of the aesthetic to the name of the variable inside aes().  Scaling is the process that assigns a unique value level of the aesthetic to a unique value.__

We can also map the variable class to show as different sizes instead of color.   
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
```
__We got a warning, because mapping an unordered variable (class) to an ordered aesthetic(size) is not a good idea.__

We can also map class to the alpha esthetic, which controls the transparency of the points.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```
__When plotting this, I got a warning signal. Warning: Using alpha for a discrete variable is not advised. ggplot2 will only use six shapes at a time. By default, additional groups will go unplotted when you use the shape aesthetic.  The x and y are aesthetics itself, the axis act as a legend.__

We can also set the aesthetic properties of the graph ourselves, for example, by making all of the points blue. 
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```
__This does not tell anything about the variables, but it allows you to change the appearance of the plot. When you manually changing the aesthetics, they go on the outside of aes(). You can change the color, the size of the points, and the shape of a point (by using a number).__

### Common Problems 
One common issue with ggplot2 is putting the + in the wrong place.  The + must come at the end of the line, not the start.

### Facets 
You can split your plot into facets, which are subplots that each display one subset of the data. To facet your plot by a single variable, you can use facet_wrap(). The first argument has to be a formula, which is ~ followed by the variable name.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```
You can also plot on the combination of two variables and add the code facet_grid().  This is also contains a formula: variable 1 ~ variable 2. 
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)
```
### Geometric Objects
Geoms are the geometrical object that a plot uses to represent data. For example, bar charts use bar geoms.  You can change the type of geom being used in your plot. 
````{r}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))
```
__Geom_point uses points and geom_smooth uses smooth representation.__

You can change the type of line that is used in a geom_smooth graph. 
```{r}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```
__Geom_smooth() separates the cars based on their drv value (drivetrain).  4=four wheel drive, f = front wheel drive, and r = rear wheel drive.__

Many geoms use a single geometric display to display multiple rows of data.  You also can se the group aesthetic to a categorical variable to draw multiple objects.
```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
              
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
    
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE
  )
```
You also can display multiple geoms in the same plot by adding different geo functions to ggplot.  Below, we are adding a point geom and smooth geom to our plot.  
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```
However, since you are looking at the same variables, there is some duplication in the code.  You can simplify the code by writing your code like this, instead of the code above: 
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()
```
The same plot is produced, just with a simpler code. 

You also can display different aesthetics in different layers.  For example, in the plot below, we are displaying the class as different colors in the point geom layer. 
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth()
```
### Statistical Transformation 
Bar charts 
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))
```
This is a simple barplot that shows the cut of the diamond and the count. It is using counts from something called stat.  You can usually use geoms and stats interchangeably.  The same plot can be produced by changing the code slightly: 
```{r}
ggplot(data = diamonds) + 
  stat_count(mapping = aes(x = cut))
```
We also can display a bar chart of proportion, rather than count, using the following code:  
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = stat(prop), group = 1))
```
We are manipulating our data using this stat function, and it is important to know what we are doing to it.  We can use stat_summary() to summarize the y values for each x value: 
```{r}
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max,
    fun = median
  )
```

### Position adjustments 
We can also color a bar chart by using the code fill: 
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, colour = cut))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut))
```
__The code colour is telling R to only have the edges of the bars colored, but the code fill is telling R to produce a plot in which the entire bar is colored.__

You can add another variable to make a stacked barplot.  In this case, every colored rectangle represents a combination of cut and clarity: 
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))
```
If you do not want a stacked barplot, but still want to look at multiple variables you can use the code identity, dodge, or fill.  
Identity is not super useful for bar plots.  
```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + 
  geom_bar(alpha = 1/5, position = "identity")
ggplot(data = diamonds, mapping = aes(x = cut, colour = clarity)) + 
  geom_bar(fill = NA, position = "identity")
```
__The bars overlap, so to see the overlap, we have to make the bars either slightly transparent (set alpha to a low number) or completely transparent (fill = NA).__

position = "fill" makes each set of stacked bars the same hieght and you can look at proportions. 
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
```
position = "dodge" places overlapping objects directly beside one another and you are able to better compare individual values. 
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```

__For scatterplots, there is another adjustment that can be made to the plot to avoid overplotting. We want to be able to see where the mass of the data is and the position = "jitter".__
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), position = "jitter")
```

### Coordinate Systems 
We can plot coordinate systems using ggplot.  The default is a Cartesian coordinate system.  One function that we can use is coord_flip() which switches the x and y axes. 
```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()
```

coord_quickmap() sets the aspect ratio correctly for maps.  

```{r}
nz <- map_data("nz")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black") +
  coord_quickmap()
```
The function coord_polar() uses polar coordinates. 
```{r}
bar <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

bar + coord_flip()
bar + coord_polar()
```

### Graphics for Communication (Chapter 28)
The best place to start when turning an exploratory graphic into a expository graphic is with good labels.  Add labels with the labs() function.  The examplle below adds a plot title. 

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs(title = "Fuel efficiency generally decreases with engine size")
```
You can also add subtitles using code subtitle and the code caption to add text at the bottom right of the plot: 
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs(
    title = "Fuel efficiency generally decreases with engine size",
    subtitle = "Two seaters (sports cars) are an exception because of their light weight",
    caption = "Data from fueleconomy.gov"
  )
```
You can change the x and y labels too, use the code labs() to change the axis and legend titles: 
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_smooth(se = FALSE) +
  labs(
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    colour = "Car type"
  )
```

### Annotations 
You can also label individual observations or groups of observations. geom_text() is similar to geom_point(), but it has an additional aesthetic: label. This code makes it possible to add textual labels to plots.  
```{r}
library('dplyr') 
library(ggplot2)
```
Needed to download the piping option. 

```{r}
best_in_class <- mpg %>%
  group_by(class) %>%
  filter(row_number(desc(hwy)) == 1)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_text(aes(label = model), data = best_in_class)
```
geom_label() which draws a rectangle behind the text, making it easier to read our plot. 
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_label(aes(label = model), data = best_in_class, nudge_y = 2, alpha = 0.5)
```

There is overlap in the labels, so we can use the ggrepel package, which adjusts labels so they do not overlap. 

```{r}
install.packages('ggrepel')
library('ggrepel')
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_point(size = 3, shape = 1, data = best_in_class) +
  ggrepel::geom_label_repel(aes(label = model), data = best_in_class)
```
I had to download the package, ggrepel.  You can also replace the legend with labels placed on the plot, this is show in the below code and plot: 
```{r}
class_avg <- mpg %>%
  group_by(class) %>%
  summarise(
    displ = median(displ),
    hwy = median(hwy)
  )
#> `summarise()` ungrouping output (override with `.groups` argument)

ggplot(mpg, aes(displ, hwy, colour = class)) +
  ggrepel::geom_label_repel(aes(label = class),
    data = class_avg,
    size = 6,
    label.size = 0,
    segment.color = NA
  ) +
  geom_point() +
  theme(legend.position = "none")
```
You can also add a single label to the plot, in the corner of the plot.  You can create a new data frame using the code (summarise()) to compute the max and min values of the variables x and y.  
```{r}
label <- mpg %>%
  summarise(
    displ = max(displ),
    hwy = max(hwy),
    label = "Increasing engine size is \nrelated to decreasing fuel economy."
  )

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_text(aes(label = label), data = label, vjust = "top", hjust = "right")
```
```{r}
label <- tibble(
  displ = Inf,
  hwy = Inf,
  label = "Increasing engine size is \nrelated to decreasing fuel economy."
)

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_text(aes(label = label), data = label, vjust = "top", hjust = "right")
```
Putting the text in the exact corner of the plot - use +Inf and -Inf: 
```{r}
label <- tibble(
  displ = Inf,
  hwy = Inf,
  label = "Increasing engine size is \nrelated to decreasing fuel economy."
)

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_text(aes(label = label), data = label, vjust = "top", hjust = "right")
```
There are other ways to annotate your plots: 
- geom_hline() and geom_vline() to add reference lines. 
- geom_rect() to draw a rectangle around points of interest. The boundaries of the rectangle are defined by aesthetics xmin, xmax, ymin, ymax.
-geom_segment() with the arrow argument to draw attention to a point with an arrow. Use aesthetics x and y to define the starting location, and xend and yend to define the end location.

### Scales 
It is important to adjust the scales if necessary, to make your plots look better. You can adjust the axis ticks and legend keys. 
The code breaks controls the position of the ticks. Labels controls the text label associated with the tick. 
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_y_continuous(breaks = seq(15, 40, by = 5))
```
You can remove the labels from an axis, you use labels = NULL. 
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_x_continuous(labels = NULL) +
  scale_y_continuous(labels = NULL)
```
__Axis and legends are known as guides as well.  Axes are used for x and y aesthetics.__

You can use breaks when you have not that many data points and want to highlight where observations occur: 
```{r}
presidential %>%
  mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id)) +
    geom_point() +
    geom_segment(aes(xend = end, yend = id)) +
    scale_x_date(NULL, breaks = presidential$start, date_labels = "'%y")
```

You can change the position of the legends as well.  To control where the legend is, you can use theme().  Theme code controls non-data parts of a plot.  theme(legend.position = ? ) controls where the legend is drawn.  If you put legend.position = none, the legend is removed.   
```{r}
base <- ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class))

base + theme(legend.position = "left")
base + theme(legend.position = "top")
base + theme(legend.position = "bottom")
base + theme(legend.position = "right") # the default
base + theme(legend.position = "none") 

```

In addition, you can make other changes to a legend by using guides() along with guide_legend() and guide_colourbar(). 
-nrow: controls the number of rows the legend uses 
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(nrow = 1, override.aes = list(size = 4)))
#> `geom_smooth()` using method = 'loess' and formula 'y ~ x'
```

### Replacing the scale 
```{r}
ggplot(diamonds, aes(carat, price)) +
  geom_bin2d()

ggplot(diamonds, aes(log10(carat), log10(price))) +
  geom_bin2d()
```
We want to change this so that the axes are now labeled with the  transformed values.  We can change this so that the transformation is not in the aesthetic mapping, but instead in the scale: 
```{r}
ggplot(diamonds, aes(carat, price)) +
  geom_bin2d() + 
  scale_x_log10() + 
  scale_y_log10()
```

You can change the color that is being used as well. 
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv))

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv)) +
  scale_colour_brewer(palette = "Set1")
```
 
 To make it even more distinguishable between the types of points, you can use shapes.  The following code changes the shape of the dots: 
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv, shape = drv)) +
  scale_colour_brewer(palette = "Set1")
```
```{r}
presidential %>%
  mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id, colour = party)) +
    geom_point() +
    geom_segment(aes(xend = end, yend = id)) +
    scale_colour_manual(values = c(Republican = "red", Democratic = "blue"))
```
If you have a predefined mapping between values and colors you can use the code scale_colour_manual() to code for specific colors. 

### Zooming 
There are three ways to control the plot limits: adjust which data is being plotted on a plot, set the limits in each of the scales, setting xlim and ylime in coord_cartesian(). You can use that last code to zoom in on a region of the plot and focus on specific data. 
```{r}
ggplot(mpg, mapping = aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth() +
  coord_cartesian(xlim = c(5, 7), ylim = c(10, 30))

mpg %>%
  filter(displ >= 5, displ <= 7, hwy >= 10, hwy <= 30) %>%
  ggplot(aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth()
```

You can set limits on individual scales, which is basically like subsetting the data.  Most useful if you want to expand the limits. 
```{r}
suv <- mpg %>% filter(class == "suv")
compact <- mpg %>% filter(class == "compact")

ggplot(suv, aes(displ, hwy, colour = drv)) +
  geom_point()

ggplot(compact, aes(displ, hwy, colour = drv)) +
  geom_point()
```

For example, the scales are all different in those above plots.  To overcome this issue, you can share scales across multiple plots with the following code: 
```{r}
x_scale <- scale_x_continuous(limits = range(mpg$displ))
y_scale <- scale_y_continuous(limits = range(mpg$hwy))
col_scale <- scale_colour_discrete(limits = unique(mpg$drv))

ggplot(suv, aes(displ, hwy, colour = drv)) +
  geom_point() +
  x_scale +
  y_scale +
  col_scale

ggplot(compact, aes(displ, hwy, colour = drv)) +
  geom_point() +
  x_scale +
  y_scale +
  col_scale
```
 
### Themes 
Customize nondata elements of your plot using themes. 

### Saving your plots 
You can save your plots out of R and into a final write up using the codes ggsave() and knitr.ggsave().  These codes save the most recent plot. 